On the probability of finding a nonphysical solution through shadowing

### 2020-10-15 comments.tex
Edits up to proof of conjugacy.
Edits up to proof of shadowing solution being almost surely quasi physical

\documentclass[12pt]{article}
\begin{document}
\section*{Section 2}
\begin{enumerate}
    \item Should we use ``stationary measure'' or ``stationary distribution'' for the SRB measure, instead of ``equilibrium distribution''? This is because physically, the dynamical systems we consider may not be in equilibrium: systems with an attractor, when ``energy'' is dissipated, may be in nonequilibrium steady states. A fluid dynamics/physics audience may assume that we are concerned only with systems that attain an equilibrium, if we call the SRB measure 
    an equilibrium measure.
    \item Another point about the SRB measure: the way they are introduced in section 2 (as the probability distribution of a long trajectory) is the best possible. When we talk about the Birkhoff ergodic theorem, it may not be accurate to introduce the SRB measure right afterwards because the existence of the SRB measure does not follow from Birkhoff's theorem. Secondly, without going into mathematical details, we can insert information that forces us to define SRB measure: that it is a special stationary measure which is physical. That is, we can say it is the particular measure, which is smooth along the attractor filaments, and this property is enough to guarantee that it is also the same as $\mu$ defined using the long trajectory. I have given this a shot by adding a few sentences but if this does not add value to the paper and/or is confusing, please remove them!
\end{enumerate}
\subsection*{Section 2.2: Lorenz map}
\begin{enumerate}
    \item To numerically find periodic orbits of the Lorenz system using the Lorenz map, we obtain the a value of $u_3$ for a point on a periodic orbit. However, there is numerical error in our reading of $u_3$ and hence in $u_1,$ $u_2$ as well. Does this make it so that the generated orbit is not periodic but a typical chaotic orbit? The effect of numerical error may be more pronounced for periodic orbits of longer period.  
    \item I am curious also about how to find local maxima in a program. Is the `scipy.find\_peaks` function a good way?
\end{enumerate} 
\subsection*{Section 3.1}
\begin{enumerate}
    \item Reading the text around the definition of the shadowing solution, 
    the reader may wonder when a true solution is called a shadowing solution.
    What does it shadow? To clarify, we can add that a solution of $\tilde{\varphi}_s$
    is a perturbed solution (a pseudoorbit) of $\varphi.$ The shadowing solution 
    is a true trajectory of $\varphi$ that stays close, forever, to the perturbed solution.
\end{enumerate}
\subsection*{Section 3.2}
\begin{enumerate}
    \item Given two nearby maps, how can we construct a conjugacy between their trajectories?
    I think there is a Newton's method for conjugacies, can we explain this with an example?
    Since finding conjugacies is not directly relevant to analyzing the physicality of shadowing, it may suffice to point to references. 
    \item Can we give intuition for $\tilde{h}_s?$ The conjugacy helpers $\xi_{s, j}$
    are bits in the binary representation of $\tilde{h}_s.$ Does this observation provide 
    any intuition for the shadowing map? Additionally, $\xi_{s, j}$ is a recursive function that codes the orbits of $\tilde{\varphi}_s.$ The coding is as follows. If an orbit 
    visits the interval $[1+s, 2]$ an even number of times upto and including time $j$,
    then, $\xi_{s,j} =0;$ otherwise, it is one. Can $\tilde{h}_s(x)$ ever be $x$ at some point? This means that all the bits ($\xi_{s,j}(x)$) must be the same as the binary representation of $x$ itself. Also, $\tilde{h}_s$
\end{enumerate}
\end{document}

>>>>

\documentclass[12pt]{article}
\begin{document}
\section*{Section 2}
\begin{enumerate}
    \item Should we use ``stationary measure'' or ``stationary distribution'' for the SRB measure, instead of ``equilibrium distribution''? This is because physically, the dynamical systems we consider may not be in equilibrium: systems with an attractor, when ``energy'' is dissipated, may be in nonequilibrium steady states. A fluid dynamics/physics audience may assume that we are concerned only with systems that attain an equilibrium, if we call the SRB measure 
    an equilibrium measure.
    \item Another point about the SRB measure: the way they are introduced in section 2 (as the probability distribution of a long trajectory) is the best possible. When we talk about the Birkhoff ergodic theorem, it may not be accurate to introduce the SRB measure right afterwards because the existence of the SRB measure does not follow from Birkhoff's theorem. Secondly, without going into mathematical details, we can insert information that forces us to define SRB measure: that it is a special stationary measure which is physical. That is, we can say it is the particular measure, which is smooth along the attractor filaments, and this property is enough to guarantee that it is also the same as $\mu$ defined using the long trajectory. I have given this a shot by adding a few sentences but if this does not add value to the paper and/or is confusing, please remove them!
\end{enumerate}
\subsection*{Section 2.2: Lorenz map}
\begin{enumerate}
    \item To numerically find periodic orbits of the Lorenz system using the Lorenz map, we obtain the a value of $u_3$ for a point on a periodic orbit. However, there is numerical error in our reading of $u_3$ and hence in $u_1,$ $u_2$ as well. Does this make it so that the generated orbit is not periodic but a typical chaotic orbit? The effect of numerical error may be more pronounced for periodic orbits of longer period.  
    \item I am curious also about how to find local maxima in a program. Is the `scipy.find\_peaks` function a good way?
\end{enumerate} 
\subsection*{Section 3.1}
\begin{enumerate}
    \item Reading the text around the definition of the shadowing solution, 
    the reader may wonder when a true solution is called a shadowing solution.
    What does it shadow? To clarify, we can add that a solution of $\tilde{\varphi}_s$
    is a perturbed solution (a pseudoorbit) of $\varphi.$ The shadowing solution 
    is a true trajectory of $\varphi$ that stays close, forever, to the perturbed solution.
\end{enumerate}
\subsection*{Section 3.2}
\begin{enumerate}
    \item Given two nearby maps, how can we construct a conjugacy between their trajectories?
    I think there is a Newton's method for conjugacies, can we explain this with an example?
    Since finding conjugacies is not directly relevant to analyzing the physicality of shadowing, it may suffice to point to references. 
    \item Can we give intuition for $\tilde{h}_s?$ The conjugacy helpers $\xi_{s, j}$
    are bits in the binary representation of $\tilde{h}_s.$ Does this observation provide 
    any intuition for the shadowing map? Additionally, $\xi_{s, j}$ is a recursive function that codes the orbits of $\tilde{\varphi}_s.$ The coding is as follows. If an orbit 
    visits the interval $[1+s, 2]$ an even number of times upto and including time $j$,
    then, $\xi_{s,j} =0;$ otherwise, it is one.
    \item This should probably be addressed in 3.1, and relates to what is shadowed by what.
    Can we add a line that says that the two trajectories that remain close are 
    $\varphi^j \tilde{h}_s x$ and $\tilde{\varphi}^j_s x$?
\end{enumerate}
\end{document}

### 2020-10-15 main.tex
Edits up to proof of shadowing solution being almost surely quasi physical

Figure \ref{fig:tilted_tent_density} shows how fractal the conjugacy map is.
This fractal conjugacy map transforms a uniform distribution in $[0,2]$ into
a fractal distribution similar to the ones plotted in Section \ref{sec:quasiphysical}.
In fact, we will show that the fractal distribution obtained by mapping a uniform
distribution under $\tilde{h}_s$ is exactly the family of distributions
shown in Figure \ref{fig:tent_quasiphysical}.  We will also show that
a physical solution of the tilted tent map (\ref{tent_tilted}) is uniformly
distributed in $[0,2]$.  Thus, for almost any physical solution of the
tilted tent map, a shadowing solution of the original tent map, obtained
through the conjugate map $\tilde{h}_s$, has a fractal distribution.
Such shadowing solution is therefore a quasi-physical solution of
the tent map.

We first show that a physical solution of the tilted tent map 
(\ref{tent_tilted}) is uniformly distributed in $[0,2]$, for
any $0\le s<1$.
According to ergodic theory, we only need to show that the uniform
distribution is stationary under the tilted tent map.
If $x$ is uniformly distributed in $[0,2]$ with a density of 0.5,
the left and right sides of the map, from $x=0$ to $x=1+s$ and
from $x=1+s$ to 2 respectively, both has constant derivatives.
The resulting density of $\tilde\varphi_s(x)$ is
therefore uniform in $[0,2]$.

We then show that the conjugate map $\tilde{h}_s$ maps a uniform
distribution to a fractal.  Here we use the closed form
(\ref{conjugate_tilted}-\ref{conjugate_tilted_helper}).
If $x$ is a random number drawn uniformly in $[0,2]$,
it has $\frac{1+s}2$ probability of being less than $1+s$.
Thus, $\xi_{s,0}(x)=0$ with probability $\frac{1+s}2$.
This means, according to Equation (\ref{conjugate_tilted}),
$\tilde{h}_s(x)<1$ with probability $\frac{1+s}2$ --
$\tilde{h}_s(x)$ is more likely to lie in the left half
of the domain $[0,2]$.
Furthermore, for $j\ge 0$, $\xi_{s,j+1}(x)=\xi_{s,j}(x)$
with probability $\frac{1+s}2$.  This probability has direct
implication, again by Equation (\ref{conjugate_tilted}),
to the binary decimal representation of $\tilde{h}_s(x)$ --
each binary digit repeat the previous one with probability
$\frac{1+s}2$.  Now we understand that for a uniformly random
$x$, $\tilde{h}_s(x)$ is exactly the kind of initial condition
we used to construct the quasi-physical solution
in Section \ref{sec:quasiphysical}, with $p=\frac{1+s}2$.

This analysis implies that the shadowing solution is
almost surely nonphysical.  Let $x_{s,i},i=0,1,\ldots$ be a solution
to the tilted tent map with an initial condition randomly chosen
in $[0,1]$.  Then with a hundred percent probability
it is a physical solution that uniformly visits the domain $[0,2]$.
Its shadowing solution $\tilde{h}_s(x_{s,i}),i=0,1,\ldots$, however,
is a quasi-physical solution, also with a hundred percent probability.
It explores $[0,2]$ at a nonuniform frequency with a fractal density function.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_lorenz_shadow_density_lorenz_zmax_15_28_2.67.json.png}
    \hspace{0.005\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/baseline_density_lorenz_zmax_15_28_2.67.json.png}
    \\
    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_lorenz_shadow_density_lorenz_zmax_10_30_2.67.json.png}
    \hspace{0.005\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/baseline_density_lorenz_zmax_10_30_2.67.json.png}
    \\
    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_lorenz_shadow_density_lorenz_zmax_10_28_3.33.json.png}
    \hspace{0.005\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/baseline_density_lorenz_zmax_10_28_3.33.json.png}
    \caption{
    }
    \label{fig:lorenz_params_scaled}
\end{figure}
----------------------------

>>>>

Figure \ref{fig:tilted_tent_density} shows how fractal the conjugacy map is.
This fractal conjugacy map transforms a uniform distribution in $[0,2]$ into
a fractal distribution similar to the ones plotted in Section \ref{sec:quasiphysical}.
In fact, we will show that the fractal distribution obtained by mapping a uniform
distribution through $\tilde{h}_s$ is exactly the family of distributions
shown in Figure \ref{fig:tent_quasiphysical}.  We will also show that
a physical solution of the tilted tent map (Eq. \ref{tent_tilted}) is uniformly
distributed in $[0,2]$.  Thus, for almost any physical solution of the
tilted tent map, a shadowing solution of the original tent map, obtained
through the conjugate map $\tilde{h}_s$, has a fractal distribution.
Such a shadowing solution is therefore a quasi-physical solution of
the tent map.

We first show that a physical solution of the tilted tent map 
(\ref{tent_tilted}) is uniformly distributed in $[0,2]$, for
any $0\le s<1$.
According to the ergodic theory and the properties of the SRB measure of 
one-dimensional chaotic systems (\cite{young}), 
we only need to show that the uniform
distribution is stationary under the tilted tent map.
The function $\tilde{\varphi}_s$ stretches an infinitesimal
region around each $x$ by the absolute value of the derivative at $x$; 
hence, a uniform density on an infinitesimal region around $\tilde{\varphi}_s(x)$ 
is reduced by the same factor. In other words, let $\tilde{\rho}(x)$ 
be the map of a uniform density on $[0,2]$, through $\tilde{\varphi}_s$. Then, 
$\tilde{\rho}(\tilde{\varphi}_s (x)) = 0.5/|\tilde \varphi_s'(x)| + 0.5/|\tilde \varphi_s'(2-x)|,$ where the two terms express the conservation of probability 
from the two pre-images of $\tilde{\varphi}_s$. Substituting the piecewise constant 
derivative of $\tilde{\varphi}_s$ on the two intervals, $[0,1+s)$ and $[1+s,2]$,
we have $\tilde{\rho}(\tilde{\varphi}_s (x)) = 0.25(1+s) + 0.25(1-s) = 0.5,$ at 
all $x.$ Hence, a uniform density of 0.5 is unaltered by mapping through 
$\tilde\varphi_s(x)$.

We now show that the conjugacy $\tilde{h}_s$ maps a uniform
distribution to a fractal.  Here we use the closed forms Eqs. \ref{conjugate_tilted}-\ref{conjugate_tilted_helper}.
If $x$ is a random number drawn uniformly in $[0,2]$,
it has $\frac{1+s}2$ probability of being less than $1+s$.
Thus, $\xi_{s,0}(x)=0$ with probability $\frac{1+s}2$.
This means, according to Eq. \ref{conjugate_tilted},
$\tilde{h}_s(x)<1$ with probability $\frac{1+s}2$:
$\tilde{h}_s(x)$ is more likely to lie in the left half
of the domain $[0,2]$.
Furthermore, for $j\ge 0$, $\xi_{s,j+1}(x)=\xi_{s,j}(x)$
with probability $\frac{1+s}2$ since each $\tilde{\varphi}^j_s(x)$ is sampled
from the uniform distribution on $[0,2]$.  This probability has direct
implication on $\tilde{h}_s(x)$ since, again by Eq. (\ref{conjugate_tilted}),
$\xi_{s,j}$ is the $j$th bit in the binary representation of $\tilde{h}_s$ --
each bit repeats the previous one with probability
$\frac{1+s}2$.  For a uniformly random
$x$, $\tilde{h}_s(x)$ is exactly the kind of initial condition
we used to construct the quasi-physical solution
in Section \ref{sec:quasiphysical}, with $p=\frac{1+s}2$.
\nisha{This is so well set up and awesome!}
Since this analysis holds for any $x$ sampled uniformly on $[0,2]$, we can conclude that
the shadowing solution is almost surely nonphysical.  Let $x_{s,i},i=0,1,\ldots$ be a solution
to the tilted tent map with an initial condition randomly chosen
in $[0,1]$.  Then with a hundred percent probability
it is a physical solution that uniformly visits the domain $[0,2]$.
Its shadowing solution $\tilde{h}_s(x_{s,i}),i=0,1,\ldots$, however,
is a quasi-physical solution, also with a hundred percent probability.
It explores $[0,2]$ at a nonuniform frequency with a fractal density function.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_density_lorenz_zmax_15_28_2.67.json.png}
    \hspace{0.005\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/baseline_density_lorenz_zmax_15_28_2.67.json.png}
    \\
    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_density_lorenz_zmax_10_30_2.67.json.png}
    \hspace{0.005\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/baseline_density_lorenz_zmax_10_30_2.67.json.png}
    \\
    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_density_lorenz_zmax_10_28_3.33.json.png}
    \hspace{0.005\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/baseline_density_lorenz_zmax_10_28_3.33.json.png}
    \caption{
    }
    \label{fig:lorenz_params_scaled}
\end{figure}
----------------------------
